---
title: Is Jayson Tatum the Best Player from the 2017 NBA Draft?
author: Kaden Ross
date: '2022-05-01'
slug: the-rookie-phenom-jayson-tatum
categories: []
tags: []
---
  
```{r}
library(tidyverse)
library(ggbeeswarm)
library(gt)
```

```{r}
allData <- read_csv("2017NBA.csv")
fiveHighPoints <- allData %>% arrange(desc(PTS))
fiveHighPoints <- slice_head(fiveHighPoints, n = 5)
ggplot() + geom_bar(data=fiveHighPoints, aes(x=reorder(Player, PTS), weight=PTS), fill = "red")+theme_dark()+
  labs(x="Total Points", y="Player", title="Who scores the basketball the best?", subtitle="Donovan Mitchell and Jayson Tatum have showcased their special ability to score at such a young age.", caption="Source: Sports Reference | By Kaden Ross") 
```
```{r}
jarret <- allData %>% filter(Player == "Jarrett Allen")
jayson <- allData %>% filter(Player == "Jayson Tatum")
labedPlayers <- rbind2(jarret, jayson)
ggplot() + 
  geom_point(data=allData, aes(x=WS, y=WS_PER_48)) +
  geom_smooth(data=allData, aes(x=WS, y=WS_PER_48), method="lm") +
  geom_text(data =labedPlayers, aes(x=WS, y=WS_PER_48, label=Player), color="black") +
  labs(x="Win Shares", y="Win Shares per 48 minutes", title="Who has made the biggest impact in their teams wins from the 2017 NBA draft?", subtitle="Jarrett Allen's defensive skillset has him at the top of the chart.", caption="Source: Sports Reference | By Kaden Ross")
```
```{r}
VORPhigh <- allData %>% arrange(desc(VORP))
VORPhigh <- slice_head(VORPhigh, n = 10)
VORPhigh %>% 
  select(Player, VORP) %>% 
  gt() %>% 
  cols_label(
    Player = "Player",
    VORP = "VORP"
  ) %>%
  tab_header(
    title = "Who has the best Plus/Minus out of this draft class?",
    subtitle = "The leaue average for value over replacement is -2."
  ) %>% tab_style(
    style = cell_text(color = "Black", weight = "bold", align = "left"),
    locations = cells_title("title")
  ) %>% tab_style(
    style = cell_text(color = "red", align = "left"),
    locations = cells_title("subtitle")
  )  %>%
  tab_source_note(
    source_note = md("**By:** Kaden Ross  |  **Source:** Sports Reference")
  )
```




# ``` {r echo=FALSE, warning=FALSE, message=FALSE}
# library(tidyverse)
# library(ggbeeswarm)
# library(gt)
# ```
# 
# ``` {r echo=FALSE, warning=FALSE, message=FALSE}
# players22 <- read_csv("players22.csv")
# 
# freshmen <- players22 %>% filter(Class == "FR")
# big10 <- players22 %>% filter(Conference == "Big Ten")
# 
# 
# ```
# 
# 
# 
# ```{r echo=FALSE, warning=FALSE, message=FALSE}
# freshmenBigTenPTS <- freshmen %>% filter(Conference == "Big Ten") %>%
#   arrange(desc(PTS))
# freshmenBigTenPTS <- slice_head(freshmenBigTenPTS, n = 5)
# ggplot() + geom_bar(data=freshmenBigTenPTS, aes(x=reorder(Player, PTS), weight=PTS), fill = "red")+theme_dark()+
# labs(x="Total Points", y="Player", title="McGowen's unnatural scoring ability", subtitle="He proved this season that he has the NBA-potential scoring ability", caption="Source: Sports Reference | By Kaden Ross") 
# 
# 
# ```
# 
# Nebraska's Mcgowens led all the Big Ten freshman in scoring and somehow he was not awarded Big Ten Freshman of the Year. Weird huh?
# 
# We all know Nebraska is not the best basketball school, but that should not take away from the fact McGowens lead all the freshman in scoring while playing in arguably the best conference in college basketball.
# 
# 
# 
# Getting valuable playing time in a power five conference is not easy, especially for an incoming freshman. Looking at minutes played, we can see how often Bryce was on the court for the whole season. 
# 
# ```{r echo=FALSE, message=FALSE, warning=FALSE}
# 
# freshmenBigTen <- freshmen %>% filter(Conference == "Big Ten") %>% filter(MP >= 100)
# bryce <- freshmen %>% filter(Player == "Bryce McGowens")
# mal <- freshmen %>% filter(Player == "Malaki Branham")
# ggplot() +
#   geom_vline(xintercept = 502) + 
#   geom_beeswarm(
#     data=freshmenBigTen, 
#     groupOnX=FALSE, 
#     aes(x=MP, y=Player), color="grey") + 
#   geom_beeswarm(
#     data=mal, 
#     groupOnX=TRUE, 
#     aes(x=MP, y=Player), color="blue") + 
#   geom_beeswarm(
#     data=bryce, 
#     groupOnX=TRUE, 
#     aes(x=MP, y=Player), color="red")  +
# labs(x="Minutes Played", y="Player", title="Nebraska's McGowens showed how long he can stay on the court", subtitle="Bryce had the most minutes played out of every Big Ten freshman showing hhis high endurance.", caption="Source: Sports Reference | By Kaden Ross") + 
#   theme_minimal() + 
#   theme(
#     plot.title = element_text(size = 16, face = "bold"),
#     axis.title = element_text(size = 8), 
#     plot.subtitle = element_text(size=10), 
#     panel.grid.minor = element_blank(),
#     plot.title.position = "plot",
#     )
# 
# ```
#  
#  
# Having a high number of minutes played shows how valuable a player was for their team, and ultimately shows how good that player is.
# 
# Malaki Branham won the Big Ten Freshman of the Year award when he wasn't on the court as much as Bryce, and did not even come close to his total points scored. 
# 
# 
# Something doesn't add up here, right?
#  
# 
# 
# 
# The worst thing a player can do in basketball is turn the ball over. Let's look at how often Bryce Bryce committed a turnover compared to Malaki Branham.
# 
# 
# 
# ```{r echo=FALSE, message=FALSE, warning=FALSE}
# bm <- rbind2(bryce, mal)
# bm %>% 
#   select(Player, `TOV%`, TOV, G) %>% 
#   gt() %>% 
#   cols_label(
#     Player = "Player",
#     `TOV` = "Turnovers",
#     `TOV%` = "Turnover Percentage",
#     G = "Games Played"
#   ) %>%
#   tab_header(
#     title = "How trustworthy is McGowens on the court?",
#     subtitle = "Notice a five percent difference is a significant amount in basketball."
#   ) %>% tab_style(
#     style = cell_text(color = "Black", weight = "bold", align = "left"),
#     locations = cells_title("title")
#   ) %>% tab_style(
#     style = cell_text(color = "red", align = "left"),
#     locations = cells_title("subtitle")
#   )  %>%
#   tab_source_note(
#     source_note = md("**By:** Kaden Ross  |  **Source:** Sports Reference")
#   )
# ```
# 
# 
# 
# The turnover percentage is how often the player commits a turnover when the ball is in their hands. Branham's turnover percentage is clearly higher, which makes it hard for a coach to keep him on the court.
# 
# Bryce manged to play four more games than Branham this season and only ended up with nine more turnovers than Branham did.  
# 
# It is very hard to believe McGowens did not win Big Ten freshman of the year despite flourishing in his first season with the Nebraska Cornhuskers. 
# 
# 
